---
title: "NBA Player Analytics"
author: "Siyuan (Larry) Chen"
output: html_document
---

# Packages

```{r}
library(tidyverse)
library(broom)
library(rstan)
library(shinystan)
library(rstanarm)
library(loo)
```

# Reading in my dataset

```{r}
nba <- read.csv("all_seasons.csv")
```

# Basic filtering of dataset

```{r}
recent <- c("2015-16", "2016-17", "2017-18", "2018-19", "2019-20")
nba <- nba %>% 
  filter(
    season %in% recent
  )

nba %>% 
  filter(
    gp < 48
  ) %>% 
  count()

nba %>% 
  count(draft_year) %>% 
  filter(
    draft_year == "Undrafted"
  )

nba <- nba %>% 
  filter(
    gp > 55
  ) %>% 
  filter(
    draft_year != "Undrafted"
  )

nba %>% 
  filter(
    country != "USA"
  ) %>% 
  count()

nba <- nba %>% 
  mutate(
    country = ifelse(country == "USA", "USA", "International")
  ) %>% 
  mutate(
    draft_number = as.numeric(draft_number)
  )
```

# Tables 1 and 2

```{r}
nba %>% 
  filter(player_name == "Stephen Curry") %>% 
  mutate(Name = player_name) %>% 
  select(Name, net_rating, season)

nba %>% 
  filter(player_name == "Giannis Antetokounmpo") %>% 
  mutate(Name = player_name) %>% 
  select(Name, net_rating, season)
```

# Basic EDA on player net rating

```{r}
nba %>% 
  arrange(desc(net_rating)) %>% 
  select(player_name, net_rating, season, gp)
```
# EDA for Numeric Variables

Scatterplots of relationship between net rating and numeric variables, only one scatterplot was chosen for the research paper.

```{r}
ggplot(nba, aes(x = age, y = net_rating)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Net Rating versus Player Age", x = "Player Age", y = "Net Rating") +
  theme_bw()
  
ggplot(nba, aes(x = ts_pct, y = net_rating)) +
  geom_point() +
  facet_grid(. ~ country) +
  geom_smooth(method = 'lm') +
  labs(title = "Net Rating versus True Shooting", x = "True Shooting", y = "Net Rating", fill = "Nationality") +
  theme_bw()

ggplot(nba, aes(x = ast_pct, y = net_rating)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Net Rating versus Assist Percentage", x = "Assist Percentage", y = "Net Rating") +
  theme_bw()

ggplot(nba, aes(x = draft_number, y = net_rating)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Net Rating versus Draft Number", x = "Draft Number", y = "Net Rating") +
  theme_bw()

ggplot(nba, aes(x = player_height, y = net_rating)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Net Rating versus Player Height", x = "Player Height", y = "Net Rating") +
  theme_bw()

ggplot(nba, aes(x = player_weight, y = net_rating)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(title = "Net Rating versus Player Weight", x = "Player Weight", y = "Net Rating") +
  theme_bw()
```

# EDA for Categorical Variables

```{r}
ggplot(nba, aes(x = country, y = net_rating)) +
  geom_boxplot()

ggplot(nba, aes(x = draft_round, y = net_rating)) +
  geom_boxplot()
```

# Initial frequentist linear model

```{r}
lm(net_rating ~ . -X -player_name -team_abbreviation -college -season -draft_year, nba)
```

```{r}
tidy_lm <- lm(net_rating ~ . -X -player_name -team_abbreviation -college -season -draft_year, nba)
tidy(tidy_lm)
```
# Backwards Elimination on initial frequentist model

```{r}
final_be <- augment(step(tidy_lm))
```

# Begin Bayesian MCMC Analysis

We mean center Age, Player_Height, Player_Weight, GP.

```{r}
nba_model <- nba %>% 
  mutate(age_cent = age - mean(age)) %>% 
  mutate(player_height_cent = player_height - mean(player_height)) %>% 
  mutate(player_weight_cent = player_weight - mean(player_weight)) %>% 
  mutate(gp_cent = gp - mean(gp)) %>% 
  select(-age, -player_height, -player_weight, -gp)

model_ready <- nba %>% 
  select(-X, -player_name, -team_abbreviation, -college, -season, -draft_year)

full_model <- stan_glm(net_rating ~ . ,
         family = gaussian(), data = model_ready, prior = cauchy(0, 2.5),
         seed = 360)
test_model <- stan_glm(net_rating ~ . -ast_pct,
         family = gaussian(), data = model_ready, prior = cauchy(0, 2.5),
         seed = 360)
loo_compare(loo(full_model), loo(test_model))

test <- summary(full_model)
```
Test_model had 15 iterations, with each one removing a different predictor variable from the model. 

# Final MCMC Generalized Linear Model

```{r}
model_fin <- model_ready %>% 
  mutate(age_cent = age - mean(age)) %>% 
  mutate(player_height_cent = player_height - mean(player_height)) %>% 
  mutate(player_weight_cent = player_weight - mean(player_weight)) %>% 
  mutate(gp_cent = gp - mean(gp)) %>% 
  select(-age, -player_height, -player_weight, -gp)

final_model <- stan_glm(net_rating ~ . -pts -reb -usg_pct,
         family = gaussian(), data = model_fin, prior = cauchy(0, 2.5),
         seed = 360)
coef(final_model)
summary(final_model)
```

# Model Diagnostics

```{r}
pp_check(final_model, plotfun = "hist", nreps = 5)
pp_check(final_model, plotfun = "stat", stat = "mean")
pp_check(final_model, plotfun = "stat_2d", stat = c("mean", "sd"))
```

# Correlation graph for additional results

```{r}
library(GGally)
model_corr <- model_fin %>% 
  select(-draft_number)
ggcorr(model_corr,
    method = c("pairwise", "spearman"),
    nbreaks = 6,
    hjust = 0.9,
    label = TRUE,
    label_size = 2,
    color = "grey50")
```

